Hydrodynamics and rheology of adhesive vesicle doublets

### 2019-01-15 vesicle_code/docs/adhesion/adhesion.tex
worked on extensional example

Figure~\ref{fig:Dec18_vesicle_shape}(a) shows the equilibrium shapes of
vesicles with three values of reduced area: $\Delta A=0.95$, $0.75$ and
$0.6$ with $H=5$ \todo[inline]{Is this really the Hamacher constant?
It's not 0.5?} and $\delta = 0.2$.  The equilibrium vesicle shape for
$\Delta A=0.95$ is the circular cap with a flatten contact region.  This
is similar to the observed shapes of two vesicles under strong adhesive
interaction
in~\cite{RamachandranAndersonLealIsraelachvili2010_Langmuir}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Adhesion of two vesicles in an extensional flow} 
\label{sec:eflow} 
The hydrodynamics of a single vesicle in an extensional flow has
revealed novel nonlinear vesicle dynamics not found for a viscous drop
\cite{KantslerSegreSteinberg2008_PRL, ZhaoShaqfeh2013_JFM,
Narsimhan2014_JFM,  DahlNarsimhanGouveia2016_SoftMatt}.  The stagnation
point (zero-velocity point) in an extensional flow is a saddle point
where two laminar streams converge along an axis and then diverge in a
different direction.  With an active control algorithm to fix the
position of the stagnation point by adjusting the streaming flow
strength with a feedback loop~\cite{Johnson-Chavarria2011_EMJ}, a
particle can be trapped at the stagnation point for long time scales to
facilitate image acquisition or other detailed measures such as PIV of
flow inside and around the particle.  To explore the application of such
a fluid trap to measure the adhesion strength between two bound vesicle
membranes, we propose the following thought experiment.

We begin from two vesicles forming a doublet under an adhesion potential
in a quiescent flow as in Section~\ref{sec:qflow}.  Once a doublet forms
and reaches an equilibrium configuration, we turn on the fluid trap with
the stagnation point placed at the center of the vesicle doublet.  For
simplification we assume that each vesicle in the doublet has the same
reduced area, membrane area, and bending stiffness.  At low extensional
rate, we expect the vesicle doublet to stay bound at the fluid trap
stagnation point.  On the other hand, the vesicle doublet may become
unstable and eventually separate at a high extensional rate.  Thus we
expect there to exist a critical extensional rate above which the
vesicle doublet cannot stay bound under a given adhesion potential.  The
dependence of the critical extensional rate on the adhesion potential
and the mechanical vesicle properties thus provides a means to probe the
physics of membrane adhesion.

\todo[inline]{Start rewriting from here}
We consider two vesicles suspended in the extensional flow $\uu =
\chi(-x,y)$, where $\chi$ is the extensional rate.  The vesicles are
initially placed symmetrically on the $x$ axis in a quiescent flow
($\chi = 0$).  Once they reach steady-state doublet, the extensional
background flow is started.  We use this example to characterize the
required extensional rate to overcome the adhesive forces of the
doublet.  In particular, we investigate the effect of the reduced area.
For all the simulations, the length of the vesicles are fixed, while the
area is changed to achieve the desired reduced area.

We observe three different vesicle configurations that depend on the
reduced area and extensional rate.  First, the doublet is not broken and
the individual vesicles long axis is parallel to the contracting
direction (Figure~\ref{fig:extensional1}).  Second, the doublet is not
broken, but it tilts to some inclination angle that is not parallel to
either of the contracting or separation directions
(Figure~\ref{fig:extensional3}).  Finally, if the extensional rate is
sufficiently large, the doublet is broken and the vesicles tend to
$(0,\pm \infty)$.

\begin{figure}[htp]
  \includegraphics[width = 0.19\textwidth]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image01.png}
  \includegraphics[width = 0.19\textwidth]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image02.png}
  \includegraphics[width = 0.19\textwidth]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image03.png}
  \includegraphics[width = 0.19\textwidth]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image04.png}
  \includegraphics[width = 0.19\textwidth]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image05.png}
  \caption{\label{fig:extensional1} Two vesicles in an extensional flow.
  The reduced area is $0.7$, the adhesion strength is $0.7$, and the
  adhesion length scale is $0.4$, extensional rate is $\chi = 0.02$.
  The doublet tilts to an inclination angle.
  }
\end{figure}

\begin{figure}[htp]
  \includegraphics[width = 0.19\textwidth]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image01.png}
  \includegraphics[width = 0.19\textwidth]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image02.png}
  \includegraphics[width = 0.19\textwidth]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image03.png}
  \includegraphics[width = 0.19\textwidth]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image04.png}
  \includegraphics[width = 0.19\textwidth]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image05.png}
  \caption{\label{fig:extensional2} Two vesicles in an extensional flow.
  The reduced area is $0.7$, the adhesion strength is $0.7$, and the
  adhesion length scale is $0.4$, extensional rate is $\chi = 0.07$.
  The doublet tilts to an inclination angle.}
\end{figure}

\begin{figure}[htp]
  \includegraphics[width = 0.19\textwidth]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image01.png}
  \includegraphics[width = 0.19\textwidth]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image02.png}
  \includegraphics[width = 0.19\textwidth]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image03.png}
  \includegraphics[width = 0.19\textwidth]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image04.png}
  \includegraphics[width = 0.19\textwidth]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image05.png}
  \caption{\label{fig:extensional3} Two vesicles in an extensional flow.
  The reduced area is $0.7$, the adhesion strength is $0.7$, and the
  adhesion length scale is $0.4$, extensional rate is $\chi = 0.1$.  The
  vesicles extensional rate has overcome the adhesive force, and the
  vesicles doublet is broken.}
\end{figure}

Whether the doublet is broken or not and the inclination angle of the
doublet depends on the extensional rate and the vesicles' reduced area.
A phase diagram of the steady-state configuration is in
Figure~\ref{fig:extensionalPhaseDiagram}.  Final configurations that
remain symmetric to the flow (Figure~\ref{fig:extensional1} are
indicated by red dots, tilt to some inclination angle
(Figure~\ref{fig:extensional2} are indicated by blue dots, and that
separate (Figure~\ref{fig:extensional3}) are indicated by green dots.

\begin{figure}[htp]
  \includegraphics[width=0.5\textwidth]{figs/extensional_adR4em1adS7em1_phaseDiagram.pdf}
  \caption{\label{fig:extensionalPhaseDiagram} Phase diagram of vesicle
  hydrodynamics in an extensional flow.  At a fixed Hamaker constant
  $H=0.7$ and adhesion length scale $\delta=0.4$, doublets may either
  remain symmetric to the flow (red marks), or tilt to some inclination
  angle (blue dots), or the extensional rate, $\chi$, may overcome the
  adhesive force (green dots).}
\end{figure}

In Figure~\ref{fig:extensionalInclinationAngle}, we plot the inclination
angle of a doublet formed by two vesicles of reduced area $0.70$ at four
different extensional rates.

\begin{figure}[htp]
  \centering
  \includegraphics[height=0.4\textwidth]{figs/adR4em1adS7em1_ra070_inclinationAngle.pdf}
  \hfill
  \includegraphics[height=0.4\textwidth]{figs/angleDefinition.pdf}
  \caption{\label{fig:extensionalInclinationAngle} {\em Left}: The
  inclination angle of a doublet formed by vesicles of reduced area
  $0.70$ submerged in an extensional flow with an extensional rate of
  $\chi$. {\em Right}: The definition of the inclination angle.  An
  inclination angle of $\pi$ indicates a doublet whose long axis is
  parallel to the converging direction, while an inclination angle of
  $\pi/2$ indicates a doublet whose short axis is parallel to the
  converging direction.}
\end{figure}

@@

\begin{figure}[htp]
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image01.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image02.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image03.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image04.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image05.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image06.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image07.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image08.png}
  \caption{\label{fig:doublet090} The formation of a doublet in a shear
  flow.  The reduced area is $0.90$, the Hamaker constant is $H=0.7$,
  the adhesion length scale is $\delta = 0.4$, and the shear rate is
  $\chi=0.5$.}
\end{figure}

>>>>

Figure~\ref{fig:Dec18_vesicle_shape}(a) shows the equilibrium shapes of
vesicles with three values of reduced area: $\Delta A=0.95$, $0.75$ and
$0.6$ with $H=5$ \todo[inline]{Is this really the Hamaker constant?
It's not 0.5?} and $\delta = 0.2$.  The equilibrium vesicle shape for
$\Delta A=0.95$ is the circular cap with a flatten contact region.  This
is similar to the observed shapes of two vesicles under strong adhesive
interaction
in~\cite{RamachandranAndersonLealIsraelachvili2010_Langmuir}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Adhesion of two vesicles in an extensional flow} 
\label{sec:eflow} 
The hydrodynamics of a single vesicle in an extensional flow has
revealed novel nonlinear vesicle dynamics not found for a viscous drop
\cite{KantslerSegreSteinberg2008_PRL, ZhaoShaqfeh2013_JFM,
Narsimhan2014_JFM,  DahlNarsimhanGouveia2016_SoftMatt}.  The stagnation
point (zero-velocity point) in an extensional flow, $\uu =\chi(-x,y) $,
is a saddle point where two laminar streams converge along an axis and
then diverge in a different direction.  With an active control algorithm
to fix the position of the stagnation point by adjusting the streaming
flow strength with a feedback loop~\cite{Johnson-Chavarria2011_EMJ}, a
particle can be trapped at the stagnation point for long time scales to
facilitate image acquisition or other detailed measures such as PIV of
flow inside and around the particle.  To explore the application of such
a fluid trap to measure the adhesion strength between two bound vesicle
membranes, we propose the following thought experiment.

We begin from two vesicles forming a doublet under an adhesion potential
in a quiescent flow as described in Section~\ref{sec:qflow}.  Once a
doublet forms and reaches an equilibrium configuration, we turn on the
fluid trap with the stagnation point placed at the center of the vesicle
doublet.  For simplification we assume that each vesicle in the doublet
has the same reduced area, membrane area, and bending stiffness.  At low
extensional rate, $\chi$, we expect the vesicle doublet to stay bound at
the fluid trap stagnation point.  On the other hand, the vesicle doublet
may become unstable and eventually separate at higher extensional rates.
Thus we expect there to exist a critical extensional rate above which
the vesicle doublet cannot stay bound under a given adhesion potential.
The dependence of the critical extensional rate on the adhesion
potential and the mechanical vesicle properties thus provides a means to
probe the physics of membrane adhesion.

We consider six pairs of vesicles with reduced areas 0.70, 0.75, 0.80,
0.85, 0.90, and 0.95, and we use extensional rates between $10^{-2}$ and
$10^{-1}$.  For all the simulations, the length of the vesicles are
fixed, while the area is changed to achieve the desired reduced area.
Since the stagnation point can be controlled in an experimental
setting~\cite{Johnson-Chavarria2011_EMJ}, we translate the vesicles at
each time step so that the stagnation point occurs exactly in the middle
of the two vesicles.  With this adjustment, the doublet can not migrate
towards $(0,\pm \infty)$; instead, for all time, each vesicle is
centered at $\pm(x_0,y_0)$ for some $(x_0,y_0)$.

As expected, we observe two different regimes---the extensional rate is
sufficiently small that the doublet structure is not broken, and the
extensional rate is sufficiently large that the doublet is broken.  In
Figure~\ref{fig:extensionalPhaseDiagram}.

\begin{figure}[htp]
  \includegraphics[width=0.5\textwidth]{figs/extensional_adR4em1adS7em1_phaseDiagram.pdf}
  \caption{\label{fig:extensionalPhaseDiagram} Phase diagram of vesicle
  hydrodynamics in an extensional flow.  At a fixed Hamaker constant
  $H=0.7$ and adhesion length scale $\delta=0.4$.  For small extensional
  rates, the doublet structure is not broken (blue marks), while at
  large extensional rates, the doublet is broken (red marks).}
\end{figure}

In Figures~\ref{fig:extensional1}--\ref{fig:extensional3}, we show
snapshots of two vesicles of reduced area $0.70$ in an extensional flow
with three different extensional rates.  Interestingly, when the
extensional rate does not overcome the adhesion rate, the doublet does
not remain parallel to the stagnation point, but instead tilts to an
inclination angle (Figures~\ref{fig:extensional1}
and~\ref{fig:extensional2}.  We include a plot of the inclination angle
as a function of time for varying extensional rates (left plot of
Figure~\ref{fig:extensionalInclinationAngle}).  The inclination angle is
defined defined in with respect to the diverging direction (right plot
of Figure~\ref{fig:extensionalInclinationAngle}).  When the doublet is
broken at reduced area $0.70$, the extensional rate is sufficiently
large to align the long axis of the vesicle with the diverging direction
and then the vesicles separate.  However, for larger reduced areas, the
separation dynamics is different.  Instead, the vesicles tilt to an
inclination angle and then separate (Figure~\ref{fig:extensional4}).

\begin{figure}[htp]
  \includegraphics[width = 0.19\textwidth]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image01.png}
  \includegraphics[width = 0.19\textwidth]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image02.png}
  \includegraphics[width = 0.19\textwidth]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image03.png}
  \includegraphics[width = 0.19\textwidth]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image04.png}
  \includegraphics[width = 0.19\textwidth]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image05.png}
  \caption{\label{fig:extensional1} Two vesicles in an extensional flow.
  The reduced area is $0.7$, the Hamaker constant is $H = 0.7$, and the
  adhesion length scale is $\delta = 0.4$, extensional rate is $\chi =
  0.02$.  The doublet tilts to an inclination angle.
  }
\end{figure}

\begin{figure}[htp]
  \includegraphics[width = 0.19\textwidth]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image01.png}
  \includegraphics[width = 0.19\textwidth]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image02.png}
  \includegraphics[width = 0.19\textwidth]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image03.png}
  \includegraphics[width = 0.19\textwidth]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image04.png}
  \includegraphics[width = 0.19\textwidth]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image05.png}
  \caption{\label{fig:extensional2} Two vesicles in an extensional flow.
  The reduced area is $0.7$, the Hamaker constant is $H = 0.7$, and the
  adhesion length scale is $\delta = 0.4$, extensional rate is $\chi =
  0.07$.  The doublet tilts to an inclination angle.}
\end{figure}

\begin{figure}[htp]
  \includegraphics[width = 0.19\textwidth]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image01.png}
  \includegraphics[width = 0.19\textwidth]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image02.png}
  \includegraphics[width = 0.19\textwidth]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image03.png}
  \includegraphics[width = 0.19\textwidth]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image04.png}
  \includegraphics[width = 0.19\textwidth]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image05.png}
  \caption{\label{fig:extensional3} Two vesicles in an extensional flow.
  The reduced area is $0.7$, the Hamaker constant is $H = 0.7$, and the
  adhesion length scale is $\delta = 0.4$, extensional rate is $\chi =
  0.1$.  The vesicles extensional rate has overcome the adhesive force,
  and the vesicles doublet is broken.}
\end{figure}

\begin{figure}[htp]
  \includegraphics[width = 0.19\textwidth]{figs/extensional_adR4em1adS7em1Chi1em1_ra080_image01.png}
  \includegraphics[width = 0.19\textwidth]{figs/extensional_adR4em1adS7em1Chi1em1_ra080_image02.png}
  \includegraphics[width = 0.19\textwidth]{figs/extensional_adR4em1adS7em1Chi1em1_ra080_image03.png}
  \includegraphics[width = 0.19\textwidth]{figs/extensional_adR4em1adS7em1Chi1em1_ra080_image04.png}
  \includegraphics[width = 0.19\textwidth]{figs/extensional_adR4em1adS7em1Chi1em1_ra080_image05.png}
  \caption{\label{fig:extensional3} Two vesicles in an extensional flow.
  The reduced area is $0.8$, the Hamaker constant is $H = 0.7$, and the
  adhesion length scale is $\delta0.4$, extensional rate is $\chi =
  0.1$.  Initially, the doublet tilts to an inclination angle, and then
  the extensional rate overcomes the adhesive force and the doublet is
  broken.}
\end{figure}

\todo[inline]{Plot of the final inclination angle of doublets that
survive the extensional rate}

\begin{figure}[htp]
  \begin{minipage}{0.6\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figs/adR4em1adS7em1_ra070_inclinationAngle.pdf}
  \end{minipage}
  \hfill
  \begin{minipage}{0.33\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figs/angleDefinition.pdf}
  \end{minipage}
  \caption{\label{fig:extensionalInclinationAngle} {\em Left}: The
  inclination angle of a doublet formed by vesicles of reduced area
  $0.70$ submerged in an extensional flow with an extensional rate of
  $\chi$. {\em Right}: The definition of the inclination angle.  An
  inclination angle of $\pi$ indicates a doublet whose long axis is
  parallel to the converging direction, while an inclination angle of
  $\pi/2$ indicates a doublet whose short axis is parallel to the
  converging direction.}
\end{figure}

@@

\begin{figure}[htp]
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image01.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image02.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image03.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image04.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image05.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image06.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image07.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image08.png}
  \caption{\label{fig:doublet090} The formation of a doublet in a shear
  flow.  The reduced area is $0.90$, the Hamaker constant is $H=0.7$,
  the adhesion length scale is $\delta = 0.4$, and the shear rate is
  $\chi=0.5$.}
  \todo[inline]{Superimpose the background velocity on the leftmost plot
  to indicate the shearing motion}
\end{figure}
